Analyzing the impact of social attributes on the success of commit integration

### 2017-02-19 Paper/MSRChallenge2017.tex
Addressed most of Claire's todo's

\section{Introduction}
\clg{detail nitpick: ``et al.'' has a paragraph after the al. (and not the et).
  This is something I write about in the blog post you all claim to have read.
  Please fix it throughout.}

For managers of an open source project, it is important to ensure that the
project consists of high quality contributions.  The decision to include code in
a project has been described as a meritocracy, where only the code is a factor
in the decision~\cite{Scacchi07}.  However, to better understand successful 
project contributions, we must take into consideration and analyze 
different social attributes from the developers creating the commits,
which can be great predictors for the success of a contribution.

@@

\textbf{Related work:}
Similar studies have shown the relationship between different attributes and 
different measures of success: pull request acceptance~\cite{gousios14, 
gousios15,tsay14icse,tsay14fse}, pull request evaluation latency~\cite{Yu15}.
Tsay et al~\cite{tsay14icse} performed a very relevant study where they 
determine that the social connection between the developer who is asking the 
pull request and the integrators is a key factor torwards acceptance of a pull 
request. Unlike this previous study, in our study we look at a broader aspect of 
social aspects and how this attributes towards the success in Travis
CI jobs.\footnote{https://docs.travis-ci.com/user/customizing-the-build} 
Gousios et al~\cite{gousios14} study the relationship between the area of the 
source code being modified by the developer requesting the changes to be pulled 
and the acceptance rate of these changes is highly correlated. Also existence of 
tests and overall code quality~\cite{gousios15} are attributes that have a 
strong relationship with the acceptance rate of a pull request. Bettenburgh et 
al~\cite{bettenburgh10} discuss the impact of social interactions with 
post-release defects, Vasilescu et al~\cite{vasilescu14} explored how different 
project characteristics (programming language, project age, etc.) affect commit 
success when integrated to the automatic build process. Yu et al~\cite{yu16} 
analyze the impact of CI usage in Github and its correlation with software 
quality. Unlike the works previously listed, our study focuses on independent social 
attributes and how these relate to the success of the integration of a commit 
pushed by the developer being described by these social attributes.

@@

\paragraph{Goal}
The goal of this study is to identify and understand relationships 
between the several different social properties that estimate the
developer's experience, and the likelihood of a developer's commit
to be considered successfully integrated in the project. We define successful
integration as passing the continuous integration build tests in the project.
\clg{I might recommend a paragraph at the top of this section defining at a high
  level the key terms, like ``technical elements'' versus ``social factors.''
  You do concretize them in the next section, but conceptually here they're a
  little bit UBD.}

@@

Therefore our main research question is presented as follows:
\newline \newline
\textbf{Do social attributes, gathered 
from the Github profiles of developers,
allow for more accurate predictions than only using technical data?}
\newline \newline

@@

\clg{Perhaps it makes sense to break this down into subsections?}

To investigate our research questions, we built a data set based on the 
attributes that consisted of both technical and social factors. For the technical factors, we collected
these from the data set provided by the MSR Challenge~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy the commit's
complexity and the build status as listed below:\clg{The listed terms don't mean
  anything on their own; they're clearly grabbed from code.  Please rename them
  to something descriptive (in English) and give a quick definition of each.
  Even better is if you give some insight as to why they might be meaningful in
  your theory (or not).}
\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item git\_diff\_src\_churn 
\item gh\_diff\_files\_added 
\item gh\_diff\_files\_deleted 
\item gh\_diff\_files\_modified 
\item gh\_diff\_tests\_added 
\item gh\_diff\_tests\_deleted 
\item gh\_diff\_src\_files 
\item gh\_sloc 
\item tr\_status 
\end{itemize}

As a result of extracting just these particular technical attributes, the new
data set is conformed by a  big number of repeated rows, which would slow down
the process of collecting the social data for each of these rows, therefore, at
this point we removed the repeated rows and continued to collect the social
attributes from the developers who pushed the commits.\clg{I'm afraid this
  sentence/paragraph isn't parsing for me.  Please rephrase, clarify, expand,
  and add precision.  Also break into multiple sentences, because it's really
  really long, which doesn't help.  To start, I think the word ``conformed'' is
  not what you mean.}

To collect the social data, we created a web scraper to extract the social data
from developer's profiles.
We used the commit hash and the project name in the original data
set provided by the MSR Challenge~\cite{msr17challenge} to locate the web page
of the commit, leveraging \textit{github.com}'s URL standard for projects and
commit hashes.
%  using the syntax
% below. 
% \newline \newline
% \textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
% \newline \newline

We scraped the commit author's
\textit{username} from a commit page. If the scrapper\clg{you mean ``scraper''
  not ``scrapper''; I fixed it once but you should fix it everywhere else.} was able to find username, the
scraper went to the user's profile information
(\textit{https://github.com/\textbf{UserName}}).

The scrapper then extracted the following items from the user's
profile.\clg{slightly better than the above for technical factors because you do
  define them in the text below but I think integrating into the itemized list
  would be better.  Also, what does each of these mean?}
\begin{itemize}
\item Repositories
\item Stars
\item Followers
\item Following
\item Contributions in the last year
\end{itemize}

Repositories represents the repositories the developer has access to. This 
includes where the developer is the repository owner,where the developer is a 
collaborator, and repositories that the developer can access through an 
organization membership.\footnote{https://developer.github.com/v3/repos/} Stars 
represents a count of the projects the developer has stared. Starring  a project
means to showing approval of the repository and creating a bookmark for later 
access.\footnote{https://help.github.com/articles/about-stars/} Followers and 
Following detail the number of developers that are following and being followed 
by the developer in question. Finally, the last attribute studied is the number 
of contributions made by the developer in the last year.

We also added one item from the original\clg{...what?} data set to the social factors:
gh\_by\_core\_team\_member.
This factor indicates if the developer making a contribution is a core member 
of that project.

We used this data to create two data sets.  One with only the technical
factors and the build status, and another with both the technical and social
factors along with the build status.\clg{This is not a complete
  sentence.}\clg{also: why do you build two data sets?}

The data set used in this study~\cite{msr17challenge} is very diverse 
and consists of commits that date from far back. The social data we 
gather is recent, gathered in January 2017. In an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the data set that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
perfect, we believe that these numbers are unlikely to have changed
significantly in the last three years.\clg{...did you spot check?}

After building the data set, we converted the\clg{which?} various Travis build statuses to
successful (passed in the\clg{which?} data set) or not successful (errored, failed, canceled
and started in the data set).  This still left a big discrepancy in the two
categories, with about a 3:1 successful to not successful ratio.\clg{I think I'm
  having ah ard time here because it's not clear which dataset we're referring
  to, but I don't understand the previous sentence, and thus don't understand
  the next sentence.} To avoid having
this discrepancy alter our results, we created a separate data set where the
successful builds were randomly selected until the number of successful builds
equaled the number of not successful builds (downsampling process).\clg{How?}  We used both the
downsampled and non-downsampled data sets in our results.\clg{Why?}

@@

\begin{table*}[ht]
\centering
\begin{tabular}{l||r|r||r|r}
  & \multicolumn{2}{|c||}{Original number of instances}  & \multicolumn{2}{|c}{Downsampled} \\
\hline
  & Technical & Technical and Social & Technical & Technical and Social \\
\hline
\textbf{Correctly Classified Instances}&  77.01\% & 81.10\%  & 64.10\% & 76.09\%\\

@@

\textbf{Kappa statistic}& 0.21 &  0.42 & 0.28  & 0.52 \\

\textbf{Mean absolute error}& 0.34 &  0.28 & 0.43 & 0.32\\

\textbf{Relative absolute error}& 88.64\% & 72.42\%  & 85.35\% & 63.07\%\\
\hline
\textbf{Increase when adding social attributes}& \multicolumn{2}{|c||}{4.09\%}  & \multicolumn{2}{|c}{11.99\%}\\

\hline
\end{tabular}
\\
\center
  \caption{ Second and third columns: Comparison of performance when running 10
    fold cross validation\clg{We've discussed this phrasing surrounding 10-fold cross
      validation.  Please fix.  What is the metric here?  How do I interpret
      these numbers?} of two decision trees. The first column after
    the first double bar uses only technical attributes to predict the Travis
    Job status, the second column augments the data set with social
    attributes. Fourth and fifth columns: \clg{this sentence is not a
      sentence/is a bit awkward.} The same intuition as the previous
    two, with the difference that the data set was downsampled to provide the
    same number of successful and unsuccessful instances.\clg{...and what is the
    takeaway from this table?} \clg{Small thing: labels should be inside the
    caption otherwise the numbers sometimes get confused.  It's a stupid latex thing.}  \label{resultsTable}} 
\end{table*}

We evaluated the predictive power of the social factors using a decision
tree~\cite{Quinlan86}.
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the Weka J48 decision tree~\cite{Weka,Quinlan1993}
with default settings. The decision trees were evaluated using ten fold cross 
validation with the data in each data set.

\clg{I killed the separation between ``evaluation'' and ``results'' because I
  don't know why it existed in the first place.}
\clg{I also wouldn't mind an enumeration of research questions; the stuff in
  design (section 2) is high level and it's not obvious where you're going to go
  at this juncture.}

@@

In the original data set, 74.43\% of the builds were successful, the rest were not
successful.  This means that a classifier which guesses successful every time
would produce a classification rate of 74.43\% if ran on the full data
set. \clg{This sentence migth be better place above, when you first discuss subsampling.}

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm. The decision tree built from
only the technical indicators has a correct classification rate of 77.01\% which
means the technical attributes only increased the prediction accuracy by about 3\%.
Adding the social data increased the classifier's accuracy from 77.01\% to
81.10\%, about a 4\% increase.\clg{It's not clear from this text that you're
  talking about the unsampled data.}

On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, when predicting a successful build in
all instances would produce 50.00\% accuracy.  After including the social data
in the data set, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data.

\clg{...upshot? Also: there are various numbers in the Table that aren't
  discussed or explained either in the caption or in these results, like, what's
  a Kappa statistic, and what do we learn from them?}

@@

Overall, we have found that social attributes which proxy experience produce a 
noticeable increase in predicting build success.  While the increase was noticeable 
with the full data set, there was a much greater increase in prediction
accuracy after performing downsampling. These results show that social factors
as the ones studied in this paper are important\clg{...maybe a strong word?} predictors when
evaluating the predictive power of build commit success.

@@

\paragraph{Internal validity:}
The researchers in this study used the dataset provided by the 
 MSRChallenge~\cite{msr17challenge} to gain useful information regarding the 
 behavior of developers when pushing commits and how well these commits integrate 
 successfully with the project. The authors have used 
 Weka,\footnote{http://www.cs.waikato.ac.nz/ml/weka/} the well known tool for 
 usage of machine learning algorithms to analyze our datasets. We are aware that, 
 as every human written tool, Weka is prone to contain errors and/or bugs; 
 nonetheless, Weka is a broadly used, open source tool that is subject to high 
 quality standards and a vast user and developer base which makes it more likely 
 to be a high quality code and to have less bugs than lesser used tools.

@@

\paragraph{External validity:} 
 To tackle the concern that the results of this study might not generalize to 
 other instances outside of the studied cases, we have used a broad and rich 
 dataset~\cite{msr17challenge} in our analysis that comprises 1,359 projects. The 
 projects used in this dataset have also been filtered to be popular (having more 
 than 10 watches on Github), and to have a history of using the continuous 
 integration tool Travis CI for more than 50 builds. All these features help to 
 build the notion that since we are building our analysis on top of a vast 
 dataset comprehending a wide diversity of projects, the threat of this analysis 
 not generalizing to other projects is mitigated.

\paragraph{Construct validity:}
 Travis job success is not the only way to analyze the success of a commit, and 
 that for different reasons the job may not be successful even if the code 
 commited does not have flaws (a flacky test case for example). Since continuous 
 integration is a best known practice in industry and the use of this methodology 
 spreads out into the open source community, we consider that cases of misuse of 
 CI such that it will lead to failing travis jobs when they should have actually 
 be successful are minimum, specially since having a solid build is one of the 
 main purposes of the methodology, and taking into account that the projects in 
 out dataset are projects with a history of using continuous integration and 
 having a certain level of popularity, this threat is minimized. We therefore 
 consider Travis job success to be a proxy 
 for success when integrating a commit into a larger project.

@@

\clg{per my email: fix the bibliography. Now.}

>>>>

\section{Introduction}
For managers of an open source project, it is important to ensure that the
project consists of high quality contributions.  The decision to include code in
a project has been described as a meritocracy, where only the code is a factor
in the decision~\cite{Scacchi07}.  However, to better understand successful 
project contributions, we must take into consideration and analyze 
different social attributes from the developers creating the commits,
which can be great predictors for the success of a contribution.

@@

\textbf{Related work:}
Similar studies have shown the relationship between different attributes and 
different measures of success: pull request acceptance~\cite{gousios14, 
gousios15,tsay14icse,tsay14fse}, pull request evaluation latency~\cite{Yu15}.
Tsay et al.~\cite{tsay14icse} performed a very relevant study where they 
determine that the social connection between the developer who is asking the 
pull request and the integrators is a key factor torwards acceptance of a pull 
request. Unlike this previous study, in our study we look at a broader aspect of 
social aspects and how this attributes towards the success in Travis
CI jobs.\footnote{https://docs.travis-ci.com/user/customizing-the-build} 
Gousios et al.~\cite{gousios14} study the relationship between the area of the 
source code being modified by the developer requesting the changes to be pulled 
and the acceptance rate of these changes is highly correlated. Also existence of 
tests and overall code quality~\cite{gousios15} are attributes that have a 
strong relationship with the acceptance rate of a pull request. Bettenburgh et 
al~\cite{bettenburgh10} discuss the impact of social interactions with 
post-release defects, Vasilescu et al.~\cite{vasilescu14} explored how different 
project characteristics (programming language, project age, etc.) affect commit 
success when integrated to the automatic build process. Yu et al.~\cite{yu16} 
analyze the impact of CI usage in Github and its correlation with software 
quality. Unlike the works previously listed, our study focuses on independent social 
attributes and how these relate to the success of the integration of a commit 
pushed by the developer being described by these social attributes.

@@

\paragraph{Goal}
The goal of this study is to identify and understand the relationship 
between the several different social properties that estimate the
developer's experience such as contributions in the last year, repositories participated, stars, etc., technical properties, such as files added and modified, and the likelihood of a developer's commit
to be considered successfully integrated in the project. We define successful
integration as passing the continuous integration build tests in the project.

@@

Therefore our main research question is presented as follows:

\textbf{Do social attributes, gathered 
from the Github profiles of developers,
allow for more accurate predictions than only using technical data?}

@@

To investigate our research questions, we built a data set based on the 
attributes that consisted of both technical and social factors.

\textbf{Technical Factors:} For the technical factors, we collected
these from the data set provided by the MSR Challenge~\cite{msr17challenge}.
We extracted only the attributes in the data set that proxy the commit's
complexity such as the size of the changes being performed as listed below:\begin{itemize}
%\item tr\_original\_commit
%\item gh\_build\_started\_at 
\item Source churn: How many lines of code were changed by the new commit in this build.
\item Files added: How many files were added by this commit.
\item Files deleted: How many files were deleted by this commit.
\item Files modified: How many files were modified by this commit.
\item Tests added: How many tests were added by this commit.
\item Tests deleted: How many tests were deleted by this commit.
\item Source files: Number of files in the new commits. 
\item Source loc: Number of executable source lines of code, in the entire repository.
\item Travis status: Status of the travis build.
\end{itemize}

\textbf{Social Factors:} To collect the social data, we created a web scraper to extract the social data
from developer's profiles.
We used the commit hash and the project name in the original data
set provided by the MSR Challenge~\cite{msr17challenge} to locate the web page
of the commit, leveraging \textit{github.com}'s URL standard for projects and
commit hashes.
%  using the syntax
% below. 
% \newline \newline
% \textit{https://github.com/\textbf{NameOfProject}/commit/\textbf{CommitHash}}
% \newline \newline

We scraped the commit author's
\textit{username} from a commit page. If the scraper was able to find username, the
scraper went to the user's profile information
(\textit{https://github.com/\textbf{UserName}}).

The scraper then extracted the following items from the user's
profile.

\begin{itemize}
\item Repositories: the repositories the developer has access to. This 
includes where the developer is the repository owner,where the developer is a 
collaborator, and repositories that the developer can access through an 
organization membership.\footnote{https://developer.github.com/v3/repos/}
\item Stars: a count of the projects the developer has stared. Starring  a project
means to showing approval of the repository and creating a bookmark for later 
access.\footnote{https://help.github.com/articles/about-stars/}
\item Followers: number of developers that that are following the author. 
\item Following: number of developers that that are being followed by the author.
\item Contributions in the last year: Number of contributions made by the author in the last year.
\end{itemize}

We also added one item from the original data set provided by the MSR Challenge to the social factors:
gh\_by\_core\_team\_member.
This factor indicates if the developer making a contribution is a core member 
of that project.

\textbf{Comparing two data sets:} We used the social data gathered to create two data sets: one with only the technical
factors and the build status, and another with both the technical and social
factors, this is to be able to compare how well can the job status be predicted by just technical data, and adding social data.

The data set used in this study~\cite{msr17challenge} is very diverse 
and consists of commits that date from far back. The social data we 
gather is recent, gathered in January 2017. In an effort to mitigate the threat of
using future result to predict previous events,
we removed the commits in the data set that were not performed within the last
three years of the data set's release (2014,2015,2016).  While this is not
perfect, we believe that these numbers are unlikely to have changed
significantly in the last three years.\clg{...did you spot check? RESPONSE: Spot check that they did belong to those years? Yeah}

Since we are interested in the success of the job status of each row, we converted the Travis build statuses to just two possible values,
successful (passed in the original provided data set) and not successful (errored, failed, canceled
and started in the original data set). At this point we had a 3:1 successful to not successful ratio in the rows of the data sets.

In the non-downsampled data set, 74.43\% of the builds were successful, the rest were not
successful.  This means that a classifier which guesses successful every time
would produce a classification rate of 74.43\% if ran on the full data
set.

To avoid this and have a 1:1 ratio successful to not successful, we created a separate data set where the
successful builds were randomly selected until the number of successful builds
equaled the number of not successful builds (downsampling process).\clg{How?}  We used both the
downsampled and non-downsampled data sets in our results to show the difference of the behavior of the decision tree when it is trained with a equal ratio than when it is trained with a non equal ratio.

@@

\begin{table*}[ht]
\centering
\begin{tabular}{l||r|r||r|r}
  & \multicolumn{2}{|c||}{Non-Downsampled}  & \multicolumn{2}{|c}{Downsampled} \\
\hline
  & Technical & Technical and Social & Technical & Technical and Social \\
\hline
\textbf{Correctly Classified Instances}&  77.01\% & 81.10\%  & 64.10\% & 76.09\%\\

@@

%\textbf{Kappa statistic}& 0.21 &  0.42 & 0.28  & 0.52 \\

%\textbf{Mean absolute error}& 0.34 &  0.28 & 0.43 & 0.32\\

%\textbf{Relative absolute error}& 88.64\% & 72.42\%  & 85.35\% & 63.07\%\\
\hline
\textbf{Increase when adding social attributes}& \multicolumn{2}{|c||}{4.09\%}  & \multicolumn{2}{|c}{11.99\%}\\

\hline
\end{tabular}
\\
\center
  \caption{Comparison of predictive power by running 10
    fold cross validation of two decision trees with a non-downsampled dataset (2nd and 3rd column) and downsampled (4th and 5th column). The higher the percentages represent the higher number of instances we are able to correctly predict. \label{resultsTable}}   
\end{table*}

We evaluated the predictive power of the social factors using a decision
tree~\cite{Quinlan86}.
Decision trees are a predictive model which uses binary decisions on the
attributes to predict a final class.  The predicted class is represented as the
leaves, while the attribute decisions used in the prediction cause the
different branches.
In this study, we used the Weka J48 decision tree~\cite{Weka,Quinlan1993}
with default settings. The decision trees were evaluated using ten fold cross 
validation with the data in each data set.
To give an answer to main research question: ``Do social attributes, gathered 
from the Github profiles of developers,
allow for more accurate predictions than only using technical data?", we present the following results:

@@

Table~\ref{resultsTable} shows the results of running the two data sets through the
default Weka J48 decision tree algorithm, first without downsampling. The decision tree built from
only the technical indicators has a correct classification rate of 77.01\% which
means the technical attributes only increased the prediction accuracy by about 3\%.
Adding the social data increased the classifier's accuracy from 77.01\% to
81.10\%, a 4.09\% increase.

On the downsampled data, the decision tree created from the technical attributes
had a classification accuracy of 64.10\%, when predicting a successful build in
all instances would produce 50.00\% accuracy.  After including the social data
in the data set, the classifier correctly predicted 76.09\% of the data set,
about a 12\% improvement from only the technical data

@@

Overall, we have found that social attributes which proxy experience produce a 
noticeable increase in predicting build success.  While the increase was noticeable 
with the full data set, there was a much greater increase in prediction
accuracy after performing downsampling. These results show that social factors
as the ones studied in this paper are significant predictors when
evaluating the predictive power of build commit success.

@@

\paragraph{Internal validity}
The researchers in this study used the dataset provided by the 
 MSRChallenge~\cite{msr17challenge} to gain useful information regarding the 
 behavior of developers when pushing commits and how well these commits integrate 
 successfully with the project. The authors have used 
 Weka,\footnote{http://www.cs.waikato.ac.nz/ml/weka/} the well known tool for 
 usage of machine learning algorithms to analyze our datasets. We are aware that, 
 as every human written tool, Weka is prone to contain errors and/or bugs; 
 nonetheless, Weka is a broadly used, open source tool that is subject to high 
 quality standards and a vast user and developer base which makes it more likely 
 to be a high quality code and to have less bugs than lesser used tools.

@@

\paragraph{External validity} 
 To tackle the concern that the results of this study might not generalize to 
 other instances outside of the studied cases, we have used a broad and rich 
 dataset~\cite{msr17challenge} in our analysis that comprises 1,359 projects. The 
 projects used in this dataset have also been filtered to be popular (having more 
 than 10 watches on Github), and to have a history of using the continuous 
 integration tool Travis CI for more than 50 builds. All these features help to 
 build the notion that since we are building our analysis on top of a vast 
 dataset comprehending a wide diversity of projects, the threat of this analysis 
 not generalizing to other projects is mitigated.

\paragraph{Construct validity}
 Travis job success is not the only way to analyze the success of a commit, and 
 that for different reasons the job may not be successful even if the code 
 commited does not have flaws (a flacky test case for example). Since continuous 
 integration is a best known practice in industry and the use of this methodology 
 spreads out into the open source community, we consider that cases of misuse of 
 CI such that it will lead to failing travis jobs when they should have actually 
 be successful are minimum, specially since having a solid build is one of the 
 main purposes of the methodology, and taking into account that the projects in 
 out dataset are projects with a history of using continuous integration and 
 having a certain level of popularity, this threat is minimized. We therefore 
 consider Travis job success to be a proxy 
 for success when integrating a commit into a larger project.

@@
